---
title: "R Programming Week 3 Assignment 2"
output: html_notebook
---
## Initialisation
```{r}
rm(list = ls())
setwd("~/OneDrive/datasciencecoursera/R_Programming/week3")
```

## Example: Caching the Mean of a Vector
In this example we introduce the <<- operator which can be used to assign a value to an object in an environment that is different from the current environment. Below are two functions that are used to create a special object that stores a numeric vector and cache's its mean.

The first function, makeVector creates a special "vector", which is really a list containing a function to

1. set the value of the vector
2. get the value of the vector
3. set the value of the mean
4. get the value of the mean
```{r}
makeVector <- function(x = numeric()) {
    m <- NULL
    set <- function(y) {
        x <<- y
        m <<- NULL
    }
    get <- function() x
    setmean <- function(mean) m <<- mean
    getmean <- function() m
    list(set = set, get = get,
         setmean = setmean,
         getmean = getmean)
}
```
The following function calculates the mean of the special "vector" created with the above function. However, it first checks to see if the mean has already been calculated. If so, it gets the mean from the cache and skips the computation. Otherwise, it calculates the mean of the data and sets the value of the mean in the cache via the setmean function.
```{r}
cachemean <- function(x, ...) {
    m <- x$getmean()
    if(!is.null(m)) {
        message("getting cached data")
        return(m)
    }
    data <- x$get()
    m <- mean(data, ...)
    x$setmean(m)
    m
}
```
Test them
```{r}
ex <- c(1, 2, 3, 4)
vect <- makeVector(ex)
cachemean(vect)
vect$getmean()
```

##Assignment: Caching the Inverse of a Matrix
Matrix inversion is usually a costly computation and there may be some benefit to caching the inverse of a matrix rather than compute it repeatedly (there are also alternatives to matrix inversion that we will not discuss here). Your assignment is to write a pair of functions that cache the inverse of a matrix.

Write the following functions:

1. makeCacheMatrix: This function creates a special "matrix" object that can cache its inverse.
2. cacheSolve: This function computes the inverse of the special "matrix" returned by makeCacheMatrix above. If the inverse has already been calculated (and the matrix has not changed), then the cachesolve should retrieve the inverse from the cache.
Computing the inverse of a square matrix can be done with the solve function in R. For example, if X is a square invertible matrix, then solve(X) returns its inverse.

For this assignment, assume that the matrix supplied is always invertible.

###makeCacheMatrix
This function creates a special "matrix" object that can cache its inverse.
```{r}
makeCacheMatrix <- function(x = matrix()) {
    # initialise matrix inverse
    inv <- NULL
    
    # set and get methods for matrix and inverse stored
    # inside this special "matrix"
    set <- function(matrix) {
        x <<- matrix
        inv <<- NULL # initially inverse is not computed
    }
    get <- function() x
    setInverse <- function(inverse) inv <<- inverse
    getInverse <- function() inv
    # return method implementions of this Cache Matrix
    list(set = set, get = get,
         setInverse = setInverse,
         getInverse = getInverse)
}
```
###cacheSolve
This function computes the inverse of the special "matrix" returned by makeCacheMatrix above. If the inverse has already been calculated (and the matrix has not changed), then the cachesolve should retrieve the inverse from the cache.
```{r}
cacheSolve <- function(x, ...) {
    # try to get already cached matrix inverse
	inv <- x$getInverse()
    if(!is.null(inv)) {
        message("getting cached data")
        return(inv)
    }
	# if it is null, do the computation:
	# get the matrix stored in the special "matrix"
	# compute its inverse
	# and cache the result
    data <- x$get()
    inv <- solve(data, ...)
    x$setInverse(inv)
    ## Return a matrix that is the inverse of 'x'
    inv
}
```
###Test them
```{r}
tc1 <- matrix(1:4, 2, 2)
mat <- makeCacheMatrix(tc1)
cacheSolve(mat)
mat$getInverse()
```


